---
title: "RWorksheet_Calvario#1"
author: "Jolien"
date: "2024-09-04"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Set up vector named age, consisting 34, 28, 22, 36, 27,18, 52, 39, 42, 29, 35, 31, 27, 22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 42, 53, 41, 51, 35, 24, 33, 41.

```{r}
age<-c(34, 28, 22, 36, 27,18, 52, 39, 42, 29, 35, 31, 27, 22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 42, 53, 41, 51, 35, 24, 33, 41)
```


```{r}
length(age)
```
# How many data points?
# 34 data points

2. 
```{r}
reciprocal <- 1/age
print(reciprocal)
```
3.
```{r}
new_age <- c(age, 0, age)
print(new_age)
```
4.
```{r}
sort (age)
```
5.

```{r}
min (age)
max (age)
```
# Find the minimum and maximum value for age.
# 17 is the minimum
# 57 is the maximum

6.
```{r}
data <- c(2.4, 2.8, 2.1, 2.5, 2.4, 2.2, 2.5, 2.3, 2.5, 2.3, 2.4, 2.7)
```

```{r}
length(data)
```
# 12 data points

7.
# the vector results in a new vector where each element is twice the original element.
```{r}
new_vector <-(data)
data * 2
```
8.1
```{r}
f <- seq(1:100)
```

8.2
```{r}
g <- seq(20:60)
```

8.3
```{r}
h <- mean(20,60)
```

8.4
```{r}
i <- sum(51,91)
```

8.5
```{r}
j <- seq(1:1000)
```

a.
```{r}
length1 <- length(f)
length2 <- length(g)
length3 <- length(h)
length4 <- length(i)
sum(length1+length2+length3+length4)
```
# 143 data points from 8.1 to 8.4

b.
```{r}
f <- seq(1:100)
g <- seq(20:60)
h <- mean(20,60)
h <- mean(20,60)
```

c.
```{r}
j <- seq(1:1000)
answer <- max(j[j <- 10])
answer
```

9. Print a vector with the integers between 1 and 100 that are not divisible by 3,5, and 7 using filter option.

```{r}
filtered_vector <- Filter(function(i) { all(i %% c(3, 5, 7) != 0) }, seq(1, 100))
filtered_vector
```
10. Generate a sequence backwards of the integers from 1 to 100.
Write the R code and its output.

```{r}
backwards_seq <- seq(100, 1)
backwards_seq
```
11. List all the natural numbers below 25 that are multiples of 3 or 5.
Find the sum of these multiples.

```{r}
multiples <- c()

for (i in 1:24) {
  if (i %% 3 == 0 || i %% 5 == 0) {
    multiples <- c(multiples, i)
  }
}

sum_multiples <- sum(multiples)

# Output the results
multiples
sum_multiples
```
# List of multiples.
# [1]  3  5  6  9 10 12 15 18 20 21 24

# Sum of multiples.
# [1] 143

a. How many data points from 10 to 11?
```{r}
num_data_points <- length(multiples)
num_data_points
```
# 11 data points

12.  Enter this statement:
x <- {0 + x + 5 + }
Describe the output.

```{r}
x <- {0 + x + 5}
```
# R will produce an error message because the expression is not syntactically correct.


13. *Set up a vector named score, consisting of 72, 86, 92, 63, 88, 89, 91, 92, 75,
75 and 77. To access individual elements of an atomic vector, one generally uses the
x[i] construction.
Find x[2] and x[3]. Write the R code and its output.


```{r}
score <- c(72, 86, 92, 63, 88, 89, 91, 92, 75, 75, 77)

second_element <- score[2]
third_element <- score[3]

second_element
third_element
```
14. # a. Create the vector

```{r}
a <- c(1, 2, NA, 4, NA, 6, 7)
a[is.na(a)] <- 999

print(a, na.print="-999")
```
15. Follow the codes below:
name = readline(prompt="Input your name: ")
age = readline(prompt="Input your age: ")
print(paste("My name is",name, "and I am",age ,"years old."))
print(R.version.string)
What is the output of the above code?

```{r}
name = readline(prompt="Input your name: ")
age = readline(prompt="Input your age: ")
print(paste("My name is", name, "and I am", age, "years old."))
print(R.version.string)
```

